---
title: "Parliamentary Money in Brazil"
date: "July 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About the data

The parliamentarians receive a certain amount of money to spend with their needs and staff. The details of with what, where, when and how much they spend is accessible freely and openly to any person. The original data may be found [here](http://www2.camara.leg.br/transparencia/cota-para-exercicio-da-atividade-parlamentar/dados-abertos-cota-parlamentar). The file is originally available in XML, but it was converted to CSV by [nazareno](https://github.com/nazareno) in order to be easily used in this article. All the data is from 2016.  

#### The file structure

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)

untar("data/ano-atual.csv.tgz", files="ano-atual.csv", exdir="temp")

source("https://github.com/nazareno/ciencia-de-dados-1/raw/master/1-EDA/problema%201/gastos_lib.R")

expenses = ler_gastos("temp/ano-atual.csv")
quota_state = read.csv("data/valor-cota-por-estado.csv")
```

There are `r nrow(expenses)` observations and `r ncol(expenses)` variables in the dataset file.  

The name of the variables are:  
```{r}
colnames(expenses)
```

The meanings of the variables may be found at the Camara's website linked at the beginning of this page.  

A little summary of the values spent may be seen below.

```{r}
summary(expenses$vlrDocumento)
```

## About the Quota

In Brazil each state has a quota per month for its parliamentarians. It is assumed they will not use more than that.

```{r fig.align="center"}
quota_bar = ggplot(quota_state) +
  geom_bar(aes(x = reorder(estado, valor), y = valor),
           stat = "identity", fill = "#64B5F6") +
  coord_flip() +
  labs(title = "Quota by State",
       y = "Amount in Reais (R$)",
       x = "State") +
  theme_minimal()

quota_bar
```

At first glance, it is interesting to see if there are parliamentarians that have spent more than their Quota. If there are, how many per state? I don't know if they need to explain why the expenses were over the quota, but it looks like it is something normal among the parliamentarians. As you can see:

```{r fig.align="center"}
parliamentary_per_party = expenses %>%
  group_by(sgUF) %>%
  summarise(parliamentaries = length(unique(nuDeputadoId)))

over_quota = expenses %>%
  left_join(quota_state, by = c("sgUF" = "estado")) %>%
  select(
    sgPartido, nuDeputadoId, txNomeParlamentar, sgUF,
    numMes, datEmissao, vlrDocumento, vlrLiquido,
    vlrRestituicao, quota = valor) %>%
  group_by(nuDeputadoId, txNomeParlamentar, sgUF, numMes, quota) %>%
  summarise(expensesMonth = sum(vlrDocumento))

f <- function(quota, expensesMonth) {
  options = c("OVER", "UNDER")
  result = options[(quota > expensesMonth) + 1]
  return(result)
}
over_quota = over_quota %>%
  mutate(over = f(quota, expensesMonth))
  

ggplot(over_quota, aes(x = sgUF)) +
  geom_bar(aes(fill = over)) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Time That The Expenses Were Under Or Over The Quota",
       y = "Times",
       x = "State",
       legend = "OVer")
```

```{r fig.align="center"}
parl_over = over_quota %>%
  group_by(nuDeputadoId) %>%
  summarise(over = min(unique(over))) %>%
  filter(!is.na(over)) %>%
  group_by(over) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count*100/sum(count), 2))

perc = ggplot(parl_over, aes(x = "", y = percentage)) +
  geom_bar(aes(fill = over, width = 1), stat = "identity") +
  coord_polar("y") +
  theme_void() +
  geom_text(aes(y = percentage/3 + c(0,cumsum(percentage)[-length(percentage)]),
                label = paste(percentage, "%")), size=rel(5)) +
  ggtitle("Parliamentarians That Spent More Than The Quota")

perc
```

You may have noticed that there is a category called **NA** in the chart **Time That The Expenses Were Under Or Over The Quota**. It appears there because there are some parliamentarians that are not registered as part of one state. More research needs to be done in order to verify if those parliamentarians are allowed to use the public's money to any of the activities, or if they have any limits to their use.
The parliamentarians that are not listed in a state are:

```{r fig.align="center"}
not_listed_state = over_quota %>%
  filter(is.na(sgUF)) %>%
  group_by(txNomeParlamentar) %>%
  summarize(totalExpenses = sum(expensesMonth)) %>%
  arrange(desc(totalExpenses)) %>%
  rename("Parliamentarian" = txNomeParlamentar,
         "Total Expenses (R$)" = totalExpenses)

kable(not_listed_state)
```

## Some questions
As a first approach, some questions should be made in relation to this data. They are: a. In which categories do the Brazilian parliament spend the most money? &  b. Which expenses are the ones that have a greater variation?
There are `r length(unique(expenses$txtDescricao))` categories of expenses in the data. Some of them look like a derivation of other ones, but for some reason they were placed separately. When this data was collected `r sum(expenses$vlrDocumento)` Reais were spent.

To answer the first Question (a) I used the following bar chart:

```{r fig.align="center", fig.width = 10}
expenses_category = expenses %>%
  group_by(txtDescricao) %>%
  summarise(total_expenses = sum(vlrDocumento))

ggplot(expenses_category,
       aes(x = reorder(txtDescricao, total_expenses), y = total_expenses / 1e+6)) +
  geom_bar(stat = "identity", fill = "#64B5F6") +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  ylab("Amout in Million of Reais (R$)") +
  ggtitle("Amount of Money by Category") +
  coord_flip()
  
```

To answer the second question (b) I used the density chart. It shows the concentration of money spent in a region.

```{r fig.height = 30, fig.align = "center"}
ggplot(expenses) +
  geom_density(kernel = "gaussian", mapping = aes(x = vlrDocumento)) +
  theme_minimal() +
  facet_wrap(~ txtDescricao, ncol = 1, scales = "free") +
  ylab("Category") +
  xlab("Amout of Money (R$)")
```

#### Conclusion
In almost all cases, the Parliamentarians didn't stick to their States's quota. This may show that there is no law enforcing any restrictions. Also, some types of expenses do have a common distribution between their values. However, in cases such as "DIVULG??O DA ATIVIDADE PARLAMENTAR", the use of the money is not the same for all the parliamentarians. 

This is a sensitive topic to talk about, especially based on the current situation of Brazil.